A fast and Low-Overhead Technique to Secure Programs Against Integer Overflows

### 2012-09-10 doc/Manuscript/CGO/paper.tex

@@

% An example of vulnerability.
Figure~\ref{fig:ex_buffer_overflow} illustrates an example of integer
overflow vulnerability.
This vulnerability allows an adversary to perform a buffer overrun attack
on a program that is apparently guarded against this type of exploit.
For simplicity, we will use the C datatype \texttt{char}, which represents
8-bit long numbers.
Examples of actual vulnerabilities, with larger primitive types, are given
by Brumley {\em et al.}~\cite{Brumley07}, Dietz {\em et al.}~\cite{Dietz12}, and
Zhang {\em et al.}~\cite{Zhang09}.
The function \texttt{read\_matrix} reads a matrix of bytes from a file.
This matrix is stored as a linear sequence of bytes in a vector \texttt{out}.
The first and second bytes in the file, e.g., \texttt{w} and \texttt{h},
determine the number of rows and columns in the matrix.
If the \texttt{char} datatype could represent arbitrarily long numbers, than this
function would be safe against buffer overruns, because it would only write data
on allocated memory.
However, if, for instance, we have that \texttt{w}$= 16$ and \texttt{h}$= 16$,
then \texttt{s}$= 0$, as pointed before.
In this case, a buffer of length zero would be allocated, and all the data
stored in the file would overwrite memory in the stack of activation records.
By carefully crafting an input file, and adversary can, in this way, overwrite
the return address of \texttt{read\_matrix}, diverting program's execution to
one of \texttt{libc}'s function, such as \texttt{telnet}, for instance.

@@

We have executed the instrumented programs of the integer benchmarks of SPEC
2006 CPU to probe the overhead imposed by our instrumentation.
These programs have been executed with their ``test" input sets.
We have not been able to run the binary that LLVM produces for SPEC's
\texttt{gcc} in our environment, even without any of our passes, due to an
incompatible \texttt{ctype.h} header.
Figure~\ref{fig:unprunedTable} shows the percentage of instructions that
we instrument, without the intervention of the range analysis.
The number of instrumented instructions is relatively low, compared to the
total number of instructions, because we only instrument six different
LLVM bitcodes, in a set of 57 opcodes, not counting intrinsics.
Figure~\ref{fig:unprunedTable} also shows how many instructions have caused
overflows.
On the average, 4.90\% of the instrumented sites have caused integer overflows.

%Analysis about the amount of instructions
\begin{figure}[htb]
\begin{center}
\begin{small}
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|}
\hline
Benchmark & \#I & \#II & \#II/\#I & \#O \\ \hline
470.lbm & 13,724 & 1,142 & 8.32\% & 0 \\ \hline
433.milc & 44,236 & 1,602 & 3.62\% & 11 \\ \hline
444.namd & 100,276 & 3,234 & 3.23\% & 12 \\ \hline
447.dealII & 1,381,408 & 36,157 & 2.62\% & 50 \\ \hline
450.soplex & 136,367 & 3,158 & 2.32\% & 13 \\ \hline
464.h264ref & 271,627 & 13,846 & 5.10\% & 167 \\ \hline
473.astar & 19,243 & 857 & 4.45\% & 0 \\ \hline
458.sjeng & 54,051 & 2,504 & 4.63\% & 68 \\ \hline
429.mcf & 4,725 & 165 & 3,49\% & 8 \\ \hline
471.omnetpp & 203,201 & 1,972 & 0.97\% & 2 \\ \hline
403.gcc & 1,419,456 & 18,669 & 1.32\% & N/A \\ \hline
445.gobmk & 308,475 & 14,129 & 4.58\% & 4 \\ \hline
462.libquantum & 16,297 & 928 & 5.69\% & 7 \\ \hline
401.bzip2 & 38,831 & 2,158 & 5.56\% & 0 \\ \hline
456.hmmer & 114,136 & 4,001 & 3.51\% & 0 \\ \hline
\multicolumn{1}{|r|}{Total (Average)} & 275.070 & 6.968 & 3,96\% &  \\ \hline
\end{tabular*}
\end{small}
\end{center}
\caption{\label{fig:unprunedTable}
Instrumentation without support of range analysis.
\#I: number of LLVM bitcode instructions in the original program.
\#II: number of instructions that have been instrumented.
\#O: number of instructions that actually overflowed in the dynamic tests.}
\end{figure}

>>>>

\setlength{\textfloatsep}{12.0pt plus 1.0pt minus 1.0pt}

@@

% An example of vulnerability.
Figure~\ref{fig:ex_buffer_overflow} illustrates an example of integer
overflow vulnerability.
This vulnerability allows an adversary to perform a buffer overrun attack
on a program that is apparently guarded against this type of exploit.
For simplicity, we will use the C datatype \texttt{char}, which represents
8-bit long numbers.
Examples of actual vulnerabilities, with larger primitive types, are given
by Brumley {\em et al.}~\cite{Brumley07}, Dietz {\em et al.}~\cite{Dietz12}, and
Zhang {\em et al.}~\cite{Zhang09}.
The function \texttt{read\_matrix} reads a matrix of bytes from a file.
This matrix is stored as a linear sequence of bytes in a vector \texttt{out}.
The first and second bytes in the file, e.g., \texttt{w} and \texttt{h},
determine the number of rows and columns in the matrix.
If the \texttt{char} datatype could represent arbitrarily long numbers, than this
function would be safe against buffer overruns, because it would only write data
on allocated memory.
However, if, for instance, we have that \texttt{w}$= 16$ and \texttt{h}$= 16$,
then \texttt{s}$= 0$, as pointed before.
In this case, a buffer of length zero would be allocated, and all the data
stored in the file would overwrite memory in the stack of activation records.
By carefully crafting an input file, an adversary can, in this way, overwrite
the return address of \texttt{read\_matrix}, diverting program's execution to
one of \texttt{libc}'s function, such as \texttt{telnet}, for instance.

@@

We have executed the instrumented programs of the integer benchmarks of SPEC
2006 CPU to probe the overhead imposed by our instrumentation.
These programs have been executed with their ``test" input sets.
We have not been able to run the binary that LLVM produces for SPEC's
\texttt{gcc} in our environment, even without any of our passes, due to an
incompatible \texttt{ctype.h} header.
In addition, we have not been able to collect the statistics regarding
the overflows occurred in SPEC's \texttt{bzip2}, because the log file
has became too large, due to the occurrence of more than 3.000.000.000 
overflows. Thus, the machine ran out of memory.
Figure~\ref{fig:unprunedTable} shows the percentage of instructions that
we instrument, without the intervention of the range analysis.
The number of instrumented instructions is relatively low, compared to the
total number of instructions, because we only instrument six different
LLVM bitcodes, in a set of 57 opcodes, not counting intrinsics.
Figure~\ref{fig:unprunedTable} also shows how many instructions have caused
overflows.
On the average, 4.90\% of the instrumented sites have caused integer overflows.

%Analysis about the amount of instructions
\begin{figure}[htb]
\begin{center}
\begin{small}
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|}
\hline
Benchmark & \#I & \#II & \#II/\#I & \#O \\ \hline
470.lbm & 13,724 & 1,142 & 8.32\% & 0 \\ \hline
433.milc & 44,236 & 1,602 & 3.62\% & 11 \\ \hline
444.namd & 100,276 & 3,234 & 3.23\% & 12 \\ \hline
447.dealII & 1,381,408 & 36,157 & 2.62\% & 50 \\ \hline
450.soplex & 136,367 & 3,158 & 2.32\% & 13 \\ \hline
464.h264ref & 271,627 & 13,846 & 5.10\% & 167 \\ \hline
473.astar & 19,243 & 857 & 4.45\% & 0 \\ \hline
458.sjeng & 54,051 & 2,504 & 4.63\% & 68 \\ \hline
429.mcf & 4,725 & 165 & 3,49\% & 8 \\ \hline
471.omnetpp & 203,201 & 1,972 & 0.97\% & 2 \\ \hline
403.gcc & 1,419,456 & 18,669 & 1.32\% & N/A \\ \hline
445.gobmk & 308,475 & 14,129 & 4.58\% & 4 \\ \hline
462.libquantum & 16,297 & 928 & 5.69\% & 7 \\ \hline
401.bzip2 & 38,831 & 2,158 & 5.56\% & N/A \\ \hline
456.hmmer & 114,136 & 4,001 & 3.51\% & 0 \\ \hline
\multicolumn{1}{|r|}{Total (Average)} & 275.070 & 6.968 & 3,96\% &  \\ \hline
\end{tabular*}
\end{small}
\end{center}
\caption{\label{fig:unprunedTable}
Instrumentation without support of range analysis.
\#I: number of LLVM bitcode instructions in the original program.
\#II: number of instructions that have been instrumented.
\#O: number of instructions that actually overflowed in the dynamic tests.}
\end{figure}
